Do LSTMs Understand the Licensing of Negative Polarity Items?

### 2021-12-18 paper.tex
Updates from Overleaf
section 3.1

% Remove the "review" option to generate the final version.
\usepackage{acl}
\usepackage{gb4e}
% Standard package includes
\usepackage{times}
\usepackage{latexsym}

@@

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{Final Paper for LING 380}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

@@

\section{Methods}
\subsection{Language Models}
\subsection{Surprisal}
We asses the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric that is defined in Wilcox et. al. (2018). The metric is defined to be the log inverse probability:
\[S(x_i) = -log_2(p_i|h_{i-1}),\]
where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is the softmax activation. The surprisal is measured in bits because we take a base two logarithm.

@@

\section{Representation of Negative Polarity Items}
Negative polarity items have a variety of properties. Most obviously, they should only be present in negative contexts. They are also immune to intervening material. One negative context can license multiple negative polarity items. Lastly, the negative item must be in the correct syntactic position to license a negative polarity item. In this section we demonstrate that LSTMs have learned these four properties of negative polarity items.

\subsection{Detection of Negative Polarity Items}
As discussed in the introduction, negative polarity items are only grammatically correct in negative contexts. The following examples demonstrate this fact because only the sentences with negative contexts (a,c) are grammatically permissible:
\begin{exe}
[]
\begin{xlist}
\ex Jane didn't buy any books.
\ex[*]{Jane did buy any books.}
\ex He hasn't ever read Moby Dick.
\ex[*]{He has ever read Moby Dick}
\end{xlist}
\end{exe}
In this subsection we hope to demonstrate that LSTMs have expectations for negative polarity items. We designed 22 sentences with negaitve polarity items each in a positive and negative context. We expect to see higher surprisal metric in all of the positive context sentences. 
\subsection{Robustness of Negative Polarity Items to Intervening Material}
Syntactic dependencies are unchanged by the presence intervening material. In the following example, the presence of a negative polarity item (anymore) is due to syntactic licensing by negation (don't); modifying the object doesn't change the sentence structure, so it has no effect on NPI licensing:
\begin{exe}
[]
\begin{xlist}
\ex We don't go to the supermarket anymore.
\ex We don't go down the block to the supermarket that is owned by Mark anymore. 
\end{xlist}
\end{exe}
We showed in the previous subsection that LSTMs have expectations for negative polarity items, so we attempt to determine whether those expectations remain over intervening material. For this subsection, we designed 10 sentences base sentences with NPI items. Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item. In all cases the intervening words modify the object of the base sentence.
\subsection{Ability to License Multiple Negative Polarity Items}
In most cases, a negative context can license multiple negative polarity items.
\begin{exe}
[]
\begin{xlist}
\ex He didn't want to go with anybody anywhere
\ex [*]{He did want to go with anybody anywhere} 
\end{xlist}
\end{exe}

\subsection{Syntactic Position of Negative Context}
Just because one section of a sentence is in a negative context does not mean that it is appropriate to use a negative polarity item in another clause. In the following example, the dependent clause is fixed as negative and the independent clause (which contains the NPI) is varied, so only the first variation is correct:
\begin{exe}
[]
\begin{xlist}
\ex Because he was not tired, he was not even sleeping
\ex [*]{Because he was not tired, he was even sleeping} 
\end{xlist}
\end{exe}
In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items. In order to do this we designed 10 base sentences. Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
\section{Document Body}

>>>>

% Remove the "review" option to generate the final version.
\usepackage{acl}
\usepackage{gb4e}
\noautomath
% Standard package includes
\usepackage{times}
\usepackage{latexsym}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{subfig}
\graphicspath{{./img/}}

@@

\title{Do LSTMs Understand the Licensing of Negative Polarity Items?}

@@

\section{Methods}
\subsection{Language Models}
\subsection{Surprisal}
We asses the ability of an LSTM to predict the distribution of negative polarity items by using the surprisal metric that is defined in Wilcox et. al. (2018). The metric is defined to be the log inverse probability:
\[S(x_i) = -\log_2 P(x_i\,|\,h_{i-1}),\]
where $x_i$ is the current word, $h_{i-1}$ is the LSTM's hidden state before $x$ and the probability is given by the softmax activation. The surprisal is measured in bits because we take a base two logarithm.

@@

\section{Representation of Negative Polarity Items}
Negative polarity items have a variety of properties. Most obviously, they should only be present in negative contexts. They are also immune to intervening material. One negative context can license multiple negative polarity items. Lastly, the negative item must be in the correct syntactic position to license a negative polarity item. In this section we demonstrate that LSTMs have learned these four properties negative polarity items.

\subsection{Detection of Negative Polarity Items}
As discussed in the introduction, negative polarity items are only grammatically correct in negative contexts. The following examples demonstrate this fact because only the sentences with negative contexts (a,c) are grammatically permissible:
\begin{exe}
\ex
\begin{xlist}
\ex Bill is not here yet.
\ex[*]{Bill is here yet.}
\ex They hired somebody without any talent.
\ex[*]{They hired somebody with any talent.}
\end{xlist}
\end{exe}
In this subsection we hope to demonstrate that LSTMs have expectations for negative polarity items. We designed 20 sentences with negaitve polarity items each in a positive and negative context. We expect to see higher surprisal metric in all of the positive context sentences.

The plots of surprisal in Figure 1 are two strong examples of the general trend that we uncovered; negative polarity items are much more surprising to the model when contained positive contexts rather than negative contexts. The negative polarity item in 19 of the 20 base sentences that we created caused higher surprisal when they were placed in a positive context.

Overall, this experiment shows that our LSTM model has at least a basic understanding of when negative polarity items should be expected.

\begin{figure}
    \centering
    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{good1} }}
    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad1} }}
    \qquad
    \subfloat[Grammatical]{{\includegraphics[width=.42\linewidth]{good2} }}
    \subfloat[Ungrammatical]{{\includegraphics[width=.42\linewidth]{bad2} }}
    \caption{}
    \label{fig:example}
\end{figure}

\subsection{Robustness of Negative Polarity Items to Intervening Material}
Syntactic dependencies are unchanged by the presence intervening material. In the following example, the presence of a negative polarity item (anymore) is due to syntactic licensing by negation (don't); modifying the object doesn't change the sentence structure, so it has no effect on NPI licensing:
\begin{exe}
\ex
\begin{xlist}
\ex We don't go to the supermarket anymore.
\ex We don't go down the block to the supermarket that is owned by Mark anymore. 
\end{xlist}
\end{exe}
We showed in the previous subsection that LSTMs have expectations for negative polarity items, so we attempt to determine whether those expectations remain over intervening material. For this subsection, we designed 10 sentences base sentences with NPI items. Each of the base sentences comes in three varieties: minimal intervening material (enough to make the sentence grammatical), 3-5 words of additional intervening material, and 6-10 additional words between the negative licensor and the polarity item. In all cases the intervening words modify the object of the base sentence.
\subsection{Ability to License Multiple Negative Polarity Items}
In most cases, a negative context can license multiple negative polarity items.
\begin{exe}
\ex
\begin{xlist}
\ex He didn't want to go with anybody anywhere
\ex [*]{He did want to go with anybody anywhere} 
\end{xlist}
\end{exe}

\subsection{Licensing of Negative Polarity Items Between Clauses}
Just because one section of a sentence is in a negative context does not mean that it is appropriate to use a negative polarity item in another clause. In the following example, the dependent clause is fixed as negative and the independent clause (which contains the NPI) is varied, so only the first variation is correct:
\begin{exe}
\ex
\begin{xlist}
\ex Because he was not tired, he was not even sleeping
\ex [*]{Because he was not tired, he was even sleeping} 
\end{xlist}
\end{exe}
In this subsection we attempt to determine whether LSTMs are capable of understanding syntactic structure with regards to negative polarity items. In order to do this we designed 10 base sentences. Each sentence begins with a negative clause, and the grammatically correct version has a second negative clause and NPI wheras the incorrect version has a second postive clause and NPI.
\section{Document Body}
